\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{mathrsfs}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage[xindy, nonumberlist, nopostdot]{glossaries}

% My Custom Macros Package
\usepackage{mymacros}

\makeindex
\makeglossaries

\longnewglossaryentry{openset}
{
  name=Open Set,
  text=open set
} {
  A set $A$ is \textbf{open} (or $\mathcal{U}$-open) iff each of its points is an interior point. Observer that a set is not open iff there exists a point in the set that is not an interior point.
  \begin{note}[Examples]
    \item[]
    \begin{itemize}
      \item
        An open interval $A=(a,b)$ is an open set, for we may choose $S_p=A$ for each $p\in A$.
      \item
        The real line $\mathbb{R}$ is oepn since any open interval $S_p$ is a subset of $\mathbb{R}$, i.e. $p\in S_p\subseteq\mathbb{R}$
      \item
        The empty set $\emptyset$ is open since there is no point in $\emptyset$ which is not an interior point.
      \item
        The closed interval $B=[a,b]$ is not an open set, for any open interval containing $a$ or $b$ must contain points outside of $B$, i.e.\ the end points $a$ and $b$ are not interior points of $B$.
      \item
        Infinite open intervals $(a,\infty)$, $(-\infty,a)$ and $(-\infty,\infty)$ are open. On the other hand, infinite closed intervals $[a,\infty)$, $(-\infty,a]$ are not open sets since $a$ is not an interior point.
    \end{itemize}
  \end{note}
  \begin{note}
    The union of any number of open sets in $\mathbb{R}$ is open and the intersection of any finite number of open sets in $\mathbb{R}$ is open. For consider the class of open intervals:
    \begin{equation*}
      \{A_n=(-\frac{1}{n},\frac{1}{n}):n\in\mathbb{N}\}\text{ i.e. }\{(-1,1),(-\frac{1}{2},\frac{1}{2}),(-\frac{1}{3},\frac{1}{3}),\dots\}
    \end{equation*}
    and the intersection, $\bigcap^\infty_{n=1}A_n=\{0\}$, is a single point which is not open.
  \end{note}
}

\longnewglossaryentry{interiorpoint}
{
  name=Interior Point,
  text=interior point,
} {
  Let $A$ be a set of real numbers. A point $p\in A$ is an \textbf{interior point} of $A$ iff $p$ belongs to some open interval $S_p$ which is contained in $A$:
  \begin{equation*}
    p\in S_p\subseteq A
  \end{equation*}
}

\longnewglossaryentry{accumulationpoint}
{
  name=Accumulation (Limit) Point,
  text=accumulation point,
} {
  Let $A\subseteq\mathbb{R}$. A point $p\in\mathbb{R}$ is an \textbf{accumulation point} or \textbf{limit point} of $A$ iff every open set $G$ containing $p$ contains a point of $A$ different from $p$, i.e.:
  \begin{equation*}
    G\text{ open, }p\in G\text{ implies }A\cap(G\setminus\{p\})\neq\emptyset
  \end{equation*}
  The set of accumulation points of $A$, denoted by $A'$, is called the \textit{derived set} of $A$.
  \begin{note}[Examples]
    \item[]
    \begin{itemize}
      \item
        Every real number $p\in\mathbb{R}$ is a limit point of $\mathbb{Q}$ since every open set contains rational numbers.
      \item
        The set of integers $\mathbb{Z}$ does not have any accumulation points, i.e.\ derived set of $\mathbb{Z}$ is $\emptyset$ (as open sets in $\mathbb{R}$ can span between integers).
      \item
        Let $A=\{\frac{1}{n}:n\in\mathbb{N}\}$, the point $0$ is an accumulation point of $A$ since any open set $G$ with $0\in G$ contains an open interval $(-a_1,a_2)\subseteq G$ with $-a_1<0<a_2$, which contains points in $A$. Note that the limit point $0$ of $A$ does not belong to $A$ and there are no other limit points, i.e. $A'=\{0\}$.
        \begin{mynl}[-1][1][0.25][-0.125][0.125][white][white][-a_1][a_2]
        \end{mynl}
    \end{itemize}
  \end{note}
  \begin{note}
    \textit{"Limit point of a set"} is not to be confused with the concept \textit{"limit of a sequence"}.
  \end{note}
  \begin{note}[Bolzano-Weierstrass Theorem]
    Let $A$ be a bounded, infinite set of real numbers. Then $A$ has at least one accumulation point. However, do note that not every set, even if it is infinite, has a limit point.
  \end{note}
}

\longnewglossaryentry{closedset}
{
  name=Closed Set,
  text=closed set
} {
  is a subset $A\subseteq \mathbb{R}$, iff its complement, $A^c$, is an open set. Alternatively, $A\subseteq \mathbb{R}$ is closed iff $A$ contains each of its points of accumulation.
  \begin{note}[Examples]
    \item[]
    \begin{itemize}
      \item
        The closed interval $[a,b]$ is a closed set since its complement $(-\infty,a)\cup(b,\infty)$, the union of two open infinite intervals, is open.
      \item
        The set $A=\{\frac{1}{n}:n\in\mathbb{N}\}$ is not closed since $0$ is a limit point of $A$ but does not belong to $A$.
      \item
        The empty set $\emptyset$ and the entire line $\mathbb{R}$ are closed sets since their complements $\mathbb{R}$ and $\emptyset$, respectively, are open sets.
      \item
        Consider the open-closed interval $A=(a,b]$. Note that $A$ is not open since $b\in A$ is not an interior point of $A$, and is not closed since $a\notin A$ but is a limit point of $A$.
    \end{itemize}
  \end{note}
  \begin{note}
    \textit{Open} and \textit{Closed} sets are not inverses of each other as sets can be neither open nor closed.
  \end{note}
}

\longnewglossaryentry{cover}
{
  name=Cover,
  text=cover
} {
  A class of sets, $\mathscr{A}=\{A_i\}$, is said to \textbf{cover} a set $A$ if $A$ is contained in the union of the members of $\mathscr{A}$, i.e. $A\subseteq\cup_iA_i$.
  \begin{note}[Heine-Borel Theorem]
    Let $A=[c,d]$ be a closed and bounded interval, and $\mathcal{G}=\{G_i:i\in I\}$ be a class of open intervals which covers $A$, i.e. $A\subseteq\cup_iG_i$. Then $\mathcal{G}$ contains a finite subclass, say $\{G{i_1},G{i_2},\dots,G{i_m}\}$, which also covers $A$, i.e.,
    \begin{equation*}
      A\subseteq G_{i_1}\subseteq G_{i_2}\subseteq\dots\subseteq G_{i_m}
    \end{equation*}
    Both conditions, closed and bounded, must be satisfied by $A$ or else this theorem is not true.
  \end{note}

  \begin{note}[Examples of Heine-Borel Theorem]
    \item[]
      \begin{itemize}
        \item
          Consider the open, bounded unit interval $A=(0,1)$. The class
          \begin{equation*}
            \mathbb{G}=\{G_n=(\frac{1}{n+2},\frac{1}{n}):n\in\mathbb{N}\}
          \end{equation*}
          of open intervals covers $A$, i.e.,
          \begin{equation*}
            A\subseteq (\frac{1}{3},1)\cup(\frac{1}{4},\frac{1}{2})\cup(\frac{1}{5},\frac{1}{3})\cup\dots
          \end{equation*}
          But the union of no finite subclass of $\mathbb{G}$ contains $A$.
        \item
          Consider the closed infinite interval $A=[1,\infty)$. The class
          \begin{equation*}
            \mathbb{G}=\{(0,2),(1,3),(2,4),\dots\}
          \end{equation*}
          of open intervals covers $A$, but no infinite subclass does.
      \end{itemize}
  \end{note}
}

\longnewglossaryentry{sequence}
{
  name=Sequence,
  text=sequence
} {
  , denoted by
  \begin{equation*}
    \langle s_1,s_2,\dots\rangle\text{, }\langle s_n:n\in\mathbb{N}\rangle\text{ or }\langle s_n\rangle
  \end{equation*}
  is a function whose domain is $\mathbb{N}=\{1,2,3,\dots\}$, i.e.\ a sequence assigns a point $s_n$ to each positive integer $n\in\mathbb{N}$. The image $s_n$ or $s(n)$ of $n\in\mathbb{N}$ is called the $n$th \textit{term} of the sequence.

  \begin{note}[Bounded]
    A sequence $\langle s_n:n\in\mathbb{N}\rangle$ is said to be \textit{bounded} if its range $\{s_n:n\in\mathbb{N}\}$ is a bounded set.
  \end{note}
  \begin{note}[Notation]
    Observe that $\langle s_n:n\in\mathbb{N}\rangle$ denotes a sequence and is a function. Whereas $\{s_n:n\in\mathbb{N}\}$ denotes the range of the sequence and is a set.
  \end{note}
}

\longnewglossaryentry{convergentsequencerealnumbers}
{
  name=Convergent Sequence In Real Numbers,
  text=convergent sequence in real numbers
} {
  is defined as:
  \begin{center}
    The sequence $\langle a_1,a_2,\dots\rangle$ of real numbers converges to $b\in\mathbb{R}$ or, equivalently, the real number $b$ is the limit of the sequence $\langle a_n:n\in\mathbb{N}\rangle$, denoted by
  \end{center}
  \begin{equation*}
    \lim_{n\rightarrow\infty}a_n=b\text{, }\lim a_n=b\text{ or }a_n\rightarrow b
  \end{equation*}
  \begin{center}
    if for every $\epsilon >0$ there exists a positive integer $n_0$ such that
  \end{center}
  \begin{equation*}
    n>n_0\text{ implies }|a_n-b|<\epsilon
  \end{equation*}
  Observe that $|a_n-b|<\epsilon$ means that $b-\epsilon<a_n<b+\epsilon$ and $a_n$ belongs to the open interval $(b-\epsilon,b+\epsilon)$ containing $b$. Thus, since each term after the $n_0$th lies inside the interval $(b-\epsilon,b+\epsilon)$, only the terms before $a_{n_0}$, and there are only a finite number of them, can lie outside the interval $(b-\epsilon,b+\epsilon)$.

  \begin{note}[Alternative Definition]
    The sequence $\langle a_n:n\in\mathbb{N}\rangle$ converges to $b$ if every open set containing $b$ contains almost all, i.e.\ all but a finite number, of the terms of the sequence.
  \end{note}
  \begin{note}[Examples]
    \item[]
    \begin{itemize}
      \item
        A constant sequence $\langle a_0,a_0,\dots\rangle$ such as $\langle -2,-2,\dots\rangle$ converges to $a_0$ since each open set containing $a_0$ contains every term of the sequence.
      \item
        Each of the sequence
        \begin{equation*}
          \langle 1,\frac{1}{2},\frac{1}{3},\frac{1}{4},\dots\rangle\text{, }\langle 1,0,\frac{1}{2},0,\frac{1}{3},0,\frac{1}{4},0,\dots\rangle\text{, }\langle 1,-\frac{1}{2},\frac{1}{3},-\frac{1}{4},\dots\rangle
        \end{equation*}
        converges to 0 since any open interval containing 0 contains almost all of the terms of each of the sequences.
      \item
        Consider the sequence $\langle\frac{1}{2},\frac{1}{4},\frac{3}{4},\frac{1}{8},\frac{7}{8},\frac{1}{16},\frac{15}{16},\dots\rangle$, i.e.\ the sequence
        \[
          a_n=
          \begin{cases*}
            \frac{1}{2^{(n+2)/2}} & if $n$ is even\\
            1-\frac{1}{2^{(n+2)/2}} & if $n$ is odd
          \end{cases*}
        \]
        Observe that any open interval containing either 0 or 1 contains an infinite number of the terms of the sequence. Neither 0 nor 1, however, is a limit of the sequence. Observe, though, that 0 and 1 are accumulation points of the \textit{range} of the sequence, that is, of the set $\{\frac{1}{2},\frac{1}{4},\frac{3}{4},\frac{1}{8},\frac{7}{8},\frac{1}{16},\frac{15}{16},\dots\}$.
    \end{itemize}
  \end{note}
}

\longnewglossaryentry{subsequence}
{
  name=Subsequence,
  text=subsequence
} {
  Consider a sequence $\langle a_1,a_2,a_3,\dots\rangle$. If $\langle i_n\rangle$ is a sequence of positive integers such that $i_1<i_2<\dots$, then
  \begin{equation*}
    \langle a_{i_1},a_{i_2},a_{i_3},\dots\rangle
  \end{equation*}
  is called a \textbf{subsequence} of $\langle a_n:n\in\mathbb{N}\rangle$.

  \begin{note}
    Every bounded sequence of real numbers contains a convergent subsequence.
  \end{note}

  \begin{note}[Examples]
    \item[]
      \begin{itemize}
        \item
          Consider the sequence $\langle a_n\rangle=\langle 1,\frac{1}{2},\frac{1}{3},\frac{1}{4},\dots\rangle$. Observe that $\langle a_n\rangle=\langle 1,\frac{1}{2},\frac{1}{4},\frac{1}{8},\dots\rangle$ is a subsequence of $\langle a_n\rangle$ but that $\langle a_n\rangle=\langle \frac{1}{2},1,\frac{1}{4},\frac{1}{3},\frac{1}{6},\frac{1}{5},\dots\rangle$ is not a subsequence of $\langle a_n\rangle$ since 1 appears before $\frac{1}{2}$ in the original sequence.
        \item
          Although the sequence $\langle\frac{1}{2},\frac{1}{4},\frac{3}{4},\frac{1}{8},\frac{7}{8},\frac{1}{16},\frac{15}{16},\dots\rangle$ does not converge, it does have a convergent subsequence such as $\langle\frac{1}{2},\frac{1}{4},\frac{1}{8},\frac{1}{16},\dots\rangle$ and $\langle\frac{1}{2},\frac{3}{4},\frac{7}{8},\frac{15}{16},\dots\rangle$. On the other hand, the sequence $\langle 1,3,5,\dots\rangle$ does not have any convergent subsequences.
      \end{itemize}
  \end{note}
}

\longnewglossaryentry{convergentsequence}
{
  name=Convergent Sequence,
  text=convergent sequence
} {
  A sequence $\langle a_1,a_2,\dots\rangle$ of points in a topological space $X$ \textbf{converges} to a point $b\in X$, or $b$ is the \textit{limit} of the sequence $\langle a_n\rangle$, denoted by
  \begin{equation*}
    \lim_{n\rightarrow\infty}a_n=b\text{, }\lim a_n = b\text{ or }a_n\rightarrow b
  \end{equation*}
  iff for each open set $G$ containing $b$ there exists a positive integer $n_0\in\mathbb{N}$ such that
  \begin{equation*}
    n>n_0\text{ implies }a_n\in G
  \end{equation*}
  that is, if $G$ contains almost all, i.e.\ all except a finite number, of the terms of the sequence.

  \begin{note}[Examples]
    \item[]
    \begin{itemize}
      \item
        Let $\langle a_1,a_2,\dots\rangle$ be a sequence of points in an indiscrete topological space $(X,\mathcal{G})$. Note that: (i) $X$ is the only open set containing any point $b\in X$; and (ii) $X$ contains every term of the sequence $\langle a_n\rangle$. Accordingly, the seqeuence $\langle a_1,a_2,\dots\rangle$ contverges to every point $b\in X$.
      \item
        Let $\langle a_1,a_2,\dots\rangle$ be a sequence of points in a discrete topological space $(X,\mathcal{D})$. Now for every point $b\in X$, the singleton set $\{b\}$ is an open set containing $b$. So, if $a_n\rightarrow b$, then the set $\{b\}$ must contain almost all of the terms of the sequence. In other wrods, the sequence $\langle a_n\rangle$ converges to a point $b\in X$ iff the sequence is of the form $\langle a_1,a_2,\dots,a_{n_0},b,b,b,\dots\rangle$
      \item
        Let $\mathcal{T}$ be a cocountable topology. A sequence $\langle a_1,a_2,\dots\rangle$ converges to $b\in X$, ($a_n\rightarrow b$), iff the sequence is eventually constant, i.e. $\langle a_1,a_2,\dots,a_{n_0},b,b,b,\dots\rangle$. Prove $\Rightarrow$, we define $A^c$ to be the set consisting of the terms in the sequence $\langle a_n\rangle$ not equal to $b$, and this set is finite. Thus $A$ is an open set in $\mathcal{T}$ also containing $b$. Since $a_n\rightarrow b$, there exists $n>n_0$ such that $a_n=b$ for all $n>n_0$, hence enventually constant. \cite{convergent_sequence_in_cocountable_topology}
    \end{itemize}
  \end{note}

}

\longnewglossaryentry{topologicalspaces}
{
  name=Topological Spaces (Topology),
  text=topological spaces
} {
  Let $X$ be a non-empty set. A class $\mathcal{T}$ of subsets of $X$ is a \textbf{topology} on $X$ iff $\mathcal{T}$ satisfies the following axioms:
  \begin{itemize}
    \item[($O_1$)]
      $X$ and $\emptyset$ belong to $\mathcal{T}$.
    \item[($O_2$)]
      The union of any number of sets in $\mathcal{T}$ belongs to $\mathcal{T}$.
    \item[($O_3$)]
      The intersection of any two sets in $\mathcal{T}$ belongs to $\mathcal{T}$.
  \end{itemize}
  The members of $\mathcal{T}$ are then called $\mathcal{T}$\textit{-open sets}, of simply \textit{open sets}, and $X$ together with $\mathcal{T}$, i.e.\ the pair $(X,\mathcal{T})$ is called a \textbf{topological space}.

  \begin{note}[Examples]
    \item[]
    \begin{itemize}
      \item
        Let $\mathcal{U}=\{\cup_i I_i|I_i\in I\}$ denote the class of all open sets of real numbers where $I=\{(a,b)|a,b\in\mathbb{R}\}$. Then $\mathcal{U}$ is a topology in $\mathbb{R}$ denoted $(\mathbb{R},\mathcal{U})$, and is also called the \textit{usual topology} on $\mathbb{R}$.
      \item
        Similarily, the class $\mathcal{U}$ of all open sets in the plane $\mathbb{R}^2$ is a topology and also called the \textit{usual topology} on $\mathbb{R}^2$.
      \item
        Let $X$ be an infinite set. We can define a \textit{Cocountable Topology} (also known as \textit{Countable Complement Topology}) by declaring the empty set to be open, and a non-empty subset $U\subseteq X$ to be open if $X\setminus U$ is countable. In this case, if $X$ is countable then the cocountable topology is just the discrete topology, as the complement of any set is countable, and thus open.
    \end{itemize}
  \end{note}
}

\title{Topology}
\author{Adam Yin}
\date{}

\begin{document}

\maketitle
A glossary reference for Topology and related terms. Definitions are from various texts which I have read \cite{general_topology}.

\clearpage

\glsaddall
\setglossarystyle{tree}
\printglossaries

\medskip

\bibliographystyle{unsrt}
\bibliography{references}
\end{document}
